Batch Bayesian Optimization via Multi-objective Acquisition Ensemble for Automated Analog Circuit Design

### 2018-02-08 paper/experiments.tex
Paper update

@@

The proposed MACE algorithm was tested using eight benchmark functions and two
real-world circuits. Four state-of-the-art parallel Bayesian optimization
methods were compared, including the BLCB algorithm, the local penalization
method with EI acquisition function, the qEI and qKG methods\footnote{We
implemented the BLCB algorithm as the available open source implementations only
allow discrete input. For the EI-LP method, the code is downloaded from
https://github.com/SheffieldML/GPyOpt. The code for qEI and qKG is
downloaded from https://github.com/wujian16/Cornell-MOE}.

@@

We tested the MACE algorithm and other parallel BO method using eight commonly used benchmark
functions, as summarized in Table~\ref{tab:summaryanalygical}.

\begin{table}[htbp]
    \centering
    \caption{Summary of the analytical benchmark functions}
    \label{tab:summaryanalygical}
    \begin{tabular}{llllllll}
        \toprule
         Function           & Dimension        & Search domain             \\ \midrule
         Branin             & 2                & $[-5,  10]\times[0, 15]$  \\
         Alpine1            & 5                & $[-10, 10]^5$             \\
         Hartmann6          & 6                & $[0,   1]^6$              \\
         Eggholder          & 2                & $[-512, 512]^2$           \\
         Ackley2            & 2                & $[-32, 32]^2$             \\
         Ackley10           & 10               & $[-32, 32]^{10}$          \\
         Rosenbrock2        & 2                & $[-5,  10]^2$             \\
         Rosenbrock10       & 10               & $[-20, 20]^{10}$          \\
        \bottomrule
    \end{tabular}
    \vskip -0.2in
\end{table}

For all functions except the two 10D functions, we set the number of initial
random sampling to $N_{init} = 20$ and the number of iterations to $N_{iter} =
45$. Batch size is set to $B = 4$, the total number of function evaluations is
$N_{init} + B \times N_{iter}$. For the 10D Ackley and 10D Rosenbrock functions, we
$N_{init} = 100$ and $N_{iter} = 175$. The experiments were repeated ten
times to average the random fluctuations.

@@

The mean performances of the tested algorithms on the benchmark functions are
plotted in Figure~\ref{fig:CovPlotBenchmark}, the statistics of the final
optimization results are listed in Table~\ref{tab:result_analytical}. As can be
seen in Figure~\ref{fig:CovPlotBenchmark} and
Table~\ref{tab:result_analytical}, when running in sequential mode, the MACE
algorithm is competitive with the LCB and EI acquisition functions. The MACE-1
algorithm gave better performances than EI-1 and LCB-1 in the Eggholder,
Branin, Hartmann6, Ackley10, and Rosenbrock10 functions. Also, the parallel
MACE (MACE-4) gave the best performances among all the tested algorithms for
six out of the eight benchmark functions, and has shown dramatic speedup
compared to the sequential MACE (MACE-1).

@@

The power amplifier shown in Figure~\ref{fig:schPA} is used to test the Bayesian optimization algorithms. The
circuit is designed using the 180nm process with 12 design parameters, the
circuit is simulated by the commercial HSPICE circuit simulator to get its performances.

@@

The mean convergence plot is given in Figure~\ref{fig:resClassE}. We can see
that the MACE outperformed the BLCB and EI-LP in both sequential and batch
mode. For the batch runs, the MACE converges fastest among the three
algorithms, while the sequential MACE has similar performance as the batch
EI-LP method.

>>>>

\begin{table}[htbp]
    \centering
    \caption{Summary of the analytical benchmark functions}
    \label{tab:summaryanalygical}
    \begin{tabular}{llllllll}
        \toprule
         Function           & Dimension        & Search domain             \\ \midrule
         Branin             & 2                & $[-5,  10]\times[0, 15]$  \\
         Alpine1            & 5                & $[-10, 10]^5$             \\
         Hartmann6          & 6                & $[0,   1]^6$              \\
         Eggholder          & 2                & $[-512, 512]^2$           \\
         Ackley2            & 2                & $[-32, 32]^2$             \\
         Ackley10           & 10               & $[-32, 32]^{10}$          \\
         Rosenbrock2        & 2                & $[-5,  10]^2$             \\
         Rosenbrock10       & 10               & $[-20, 20]^{10}$          \\
        \bottomrule
    \end{tabular}
\end{table}

@@

The proposed MACE algorithm was tested using eight benchmark functions and two
real-world circuits. Four state-of-the-art parallel Bayesian optimization
methods were compared, including the BLCB algorithm, the local penalization
method with EI acquisition function (EI-LP), the qEI and qKG
methods\footnote{We implemented the BLCB algorithm as the available open source
implementations only allow discrete input. For the EI-LP method, the code is
downloaded from https://github.com/SheffieldML/GPyOpt. The code for qEI and qKG
is downloaded from https://github.com/wujian16/Cornell-MOE}. 
% For the MACE, BLCB, and EI-LP method, the ARD squared-exponential kernel is
% used and the GP models are fitted by maximum likelihood estimations(MLE); for
% the qKG and qEI methods, the Matern52 kernel is used, and the GP
% hyperparameters are integrated via MCMC sampling. The Matern52 kernel and MCMC
% integration are the default strategies of the qKG and qEI implementations and
% it is unclear in the documentation about how to change the GP settings.

@@

We tested the MACE algorithm and other parallel BO methods using eight commonly used benchmark
functions, as summarized in Table~\ref{tab:summaryanalygical}.

For all functions except the two 10D functions, we set the number of initial
random sampling to $N_{init} = 20$ and the number of iterations to $N_{iter} =
45$. Batch size is set to $B = 4$, the total number of function evaluations is
$N_{init} + B \times N_{iter}$. For the 10D Ackley and 10D Rosenbrock functions, we set
$N_{init} = 100$ and $N_{iter} = 175$. The experiments were repeated ten
times to average the random fluctuations.

@@

The mean convergence plots of the tested algorithms on the benchmark functions are
given in Figure~\ref{fig:CovPlotBenchmark}, the statistics of the final
optimization results are listed in Table~\ref{tab:result_analytical}. As can be
seen in Figure~\ref{fig:CovPlotBenchmark} and
Table~\ref{tab:result_analytical}, when running in sequential mode (MACE-1 in Table~\ref{tab:result_analytical} and Figure~\ref{fig:CovPlotBenchmark}), the MACE
algorithm is competitive with the LCB and EI acquisition functions. The MACE-1
algorithm gave better performances than EI-1 and LCB-1 in the Eggholder,
Branin, Hartmann6, Ackley10, and Rosenbrock10 functions. Also, the parallel
MACE (MACE-4) gave the best performances among all the tested algorithms for
six out of the eight benchmark functions, and has shown dramatic speedup
compared to the sequential MACE.

@@

The class-E power amplifier shown in Figure~\ref{fig:schPA} is used to test the Bayesian optimization algorithms. The
circuit is designed using the 180nm process with 12 design parameters, the
circuit is simulated by the commercial HSPICE circuit simulator to get its performances.

@@

The mean convergence plot is given in Figure~\ref{fig:resClassE}. We can see
that the MACE outperformed the BLCB and EI-LP in both sequential and batch
mode. For the batch runs, the MACE converges fastest among the three
algorithms, while the sequential MACE (MACE-1) has comparable performance as
the batch EI-LP (EI-LP-4) method.

### 2018-02-08 paper/mace.tex
Paper update

Unlike single-objective optimization, there are multiple objectives to optimize in multi-objective optimization problems\cite{MO_overview}, the multi-objective optimization problem is formulated as
\begin{equation}
    \label{eq:MOFormulation}
    \begin{aligned}
        & \text{minimize} & & f_1(\bm{x}),~\dots~,f_m(\bm{x}).
    \end{aligned}
\end{equation}
The multiple objectives to be optimizes can be conflicting, so that it is usually impossible to find a single solution that is the optimum of all objectives, the goal of multi-objective optimization algorithms is to approximate the \emph{Pareto front} of the objectives. A solution $\bm{x}_1$ is said to \emph{dominate} $\bm{x}_2$ if $\forall i \in \{1\dots m\},~f_i(\bm{x}_1) \le f_i(\bm{x}_2)$ and $\exists j \in \{1\dots m\}, f_j(\bm{x}_1) < f_j(\bm{x}_2)$. A design is \emph{Pareto-optimal} if it is not dominated by any other point in the design space and dominates at least one point. The whole set of the Pareto-optimal points in the design space is called the \emph{Pareto set}, and the set of Pareto-optimal points in the objective space is called the \emph{Pareto front}. It is often unlikely to get the whole Pareto front as there might be infinite points on the front, the goal of multi-objective optimization is to find a set of designs that approximates the true Pareto front.

@@

\begin{algorithm}
    \caption{Multi-objective Acquisition Ensemble Algorithm}
    \label{alg:MACE}
    \begin{algorithmic}[1]
        \REQUIRE Number of initial sampling points $N_{init}$, number of iterations $N_{iter}$, batch size $B$.
        \STATE Randomly sample $N_{init}$ points in the design space
        \STATE Construct initial GP model
        \FOR{t = 1, 2, \dots, $N_{iter}$}
        \STATE Construct the LCB, EI and PI functions according to \eqref{eq:LCB} and \eqref{eq:PI_EI}
        \STATE Find the Pareto front of LCB, EI, PI function using the DEMO algorithm
        \STATE Randomly sample $B$ points $\bm{x}_1, \dots, \bm{x}_B$ from the Pareto front
        \STATE Evaluate $\bm{x}_1, \dots, \bm{x}_B$ to get $y_1 = f(\bm{x}_1),~\dots~,y_B = f(\bm{x}_B)$
        \STATE Update the GP model
        \ENDFOR
        \STATE \textbf{Return} best $f(\bm{x})$ recorded during iterations
    \end{algorithmic}
\end{algorithm}

>>>>

Unlike single-objective optimization, there are multiple objectives to optimize in multi-objective optimization problems\cite{MO_overview}, the multi-objective optimization problem is formulated as
\begin{equation}
    \label{eq:MOFormulation}
    \begin{aligned}
        & \text{minimize} & & f_1(\bm{x}),~\dots~,f_m(\bm{x}).
    \end{aligned}
\end{equation}
The multiple objectives to be optimized can be conflicting so that it is usually impossible to find a single solution that is the optimum of all objectives. The goal of multi-objective optimization algorithms is to approximate the \emph{Pareto front} of the objectives. A solution $\bm{x}_1$ is said to \emph{dominate} $\bm{x}_2$ if $\forall i \in \{1\dots m\},~f_i(\bm{x}_1) \le f_i(\bm{x}_2)$ and $\exists j \in \{1\dots m\}, f_j(\bm{x}_1) < f_j(\bm{x}_2)$. A design is \emph{Pareto-optimal} if it is not dominated by any other point in the design space and dominates at least one point. The whole set of the Pareto-optimal points in the design space is called the \emph{Pareto set}, and the set of Pareto-optimal points in the objective space is called the \emph{Pareto front}. It is often unlikely to get the whole Pareto front as there might be infinite points on the Paret front, multi-objective optimization algorithms try to find a set of evenly distributed solutions that approximate the true Pareto front.

@@

\begin{algorithm}
    \caption{Multi-objective Acquisition Ensemble Algorithm}
    \label{alg:MACE}
    \begin{algorithmic}[1]
        \REQUIRE Number of initial sampling points $N_{init}$, number of iterations $N_{iter}$, batch size $B$.
        \STATE Randomly sample $N_{init}$ points in the design space
        \STATE Construct initial GP model
        \FOR{t = 1, 2, \dots, $N_{iter}$}
        \STATE Construct the LCB, EI and PI functions according to \eqref{eq:LCB} and \eqref{eq:PI_EI}
        \STATE Find the Pareto front of LCB, EI, PI functions using the DEMO algorithm
        \STATE Randomly sample $B$ points $\bm{x}_1, \dots, \bm{x}_B$ from the Pareto-optimal points
        \STATE Evaluate $\bm{x}_1, \dots, \bm{x}_B$ to get $y_1 = f(\bm{x}_1),~\dots~,y_B = f(\bm{x}_B)$
        \STATE Update the GP model
        \ENDFOR
        \STATE \textbf{Return} best $f(\bm{x})$ recorded during iterations
    \end{algorithmic}
\end{algorithm}
